Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain

### 2018-04-01 pkg/disttree/inst/draft/draft.Rnw
reran all calculations and saved all results, adapted replication code and calculations within draft.Rnw, replaced data for background of map by data from CC-BY, added corresponding reference
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2110 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 2e17c4e8585e34e07908172e0958318f66e90739

<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)
library("disttree")
library("crch")
library("countreg")
library("gamlss")
library("gamlss.cens")
gen.cens(NO, type = "left")

@@

<<echo=FALSE, results=hide>>=
#### Axams prediction 24 - 4
if(file.exists("rain_Axams_pred_24to4.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/rain_Axams_pred_24to4.rda")
  load("rain_Axams_pred_24to4.rda")
} else {
  #source("~/svn/partykit/pkg/disttree/inst/draft/rain_axams_pred_24to4.R")
  source("rain_axams_pred_24to4.R")
  library("gamlss.cens")
  gen.cens("NO", type = "left")
  res <- rain_axams_pred(seedconst = 7, ntree = 100,
                         forest_mtry = 27,   # if frac == TRUE: nvar/3 = 30
                         gamboost_cvr = TRUE,
                         frac = FALSE)

  save(res, file = "rain_Axams_pred_24to4.rda")
}

@@

pdays <- if(pday<19) c(pday, pday + 31, pday + 62, pday + 92) else c(pday, pday + 31, pday + 61, pday + 92)
pdf <- predict(res$df, newdata = res$testdata[pdays,], type = "parameter")
df_mu <- pdf$mu
df_sigma <- pdf$sigma
df_exp <- pnorm(df_mu/df_sigma) * (df_mu + df_sigma * (dnorm(df_mu/df_sigma) / pnorm(df_mu/df_sigma)))
cbind(df_exp, res$testdata[pdays,"robs"])

# plot predicted distributions together with observations
#set.seed(res$call$seedconst)
set.seed(7)
x <- c(0.01, sort(runif(500,0.01,8)))
y1 <- crch::dcnorm(x, mean = df_mu[1], sd = df_sigma[1], left = 0)
y2 <- crch::dcnorm(x, mean = df_mu[2], sd = df_sigma[2], left = 0)
y3 <- crch::dcnorm(x, mean = df_mu[3], sd = df_sigma[3], left = 0)
y4 <- crch::dcnorm(x, mean = df_mu[4], sd = df_sigma[4], left = 0)
dayending <- if(pday > 3) "th" else switch(pday, "1" = {dayending <- "st"}, "2" = {dayending <- "nd"}, "3" = {dayending <- "rd"})

@@

# predictions
pred1 <- c(res$testdata[pdays,"robs"][1], crch::dcnorm(res$testdata[pdays,"robs"][1], mean = df_mu[1], sd = df_sigma[1], left = 0))
pred2 <- c(res$testdata[pdays,"robs"][2], crch::dcnorm(res$testdata[pdays,"robs"][2], mean = df_mu[2], sd = df_sigma[2], left = 0))
pred3 <- c(res$testdata[pdays,"robs"][3], crch::dcnorm(res$testdata[pdays,"robs"][3], mean = df_mu[3], sd = df_sigma[3], left = 0))
pred4 <- c(res$testdata[pdays,"robs"][4], crch::dcnorm(res$testdata[pdays,"robs"][4], mean = df_mu[4], sd = df_sigma[4], left = 0))

@@

## PIT histograms and qqrplots

## get predicted parameter
# for learndata and testdata
{
  set.seed(7)
  ## disttree
  if(FALSE){
    # in sample
    pdt <- predict(res$dt, type = "parameter")
    dt_mu_l <- pdt$mu 
    dt_sigma_l <- pdt$sigma 
    pit_dt_l <- cbind(0, pnorm(res$learndata[,"robs"], mean = dt_mu_l, sd = dt_sigma_l))
    pit_dt_l[res$learndata[,"robs"]>0, 1] <- pit_dt_l[res$learndata[,"robs"]>0, 2]
  }  
  # out of sample
  pdt <- predict(res$dt, type = "parameter", newdata = res$testdata)
  dt_mu_t <- pdt$mu 
  dt_sigma_t <- pdt$sigma 
  pit_dt_t <- cbind(0, pnorm(res$testdata[,"robs"], mean = dt_mu_t, sd = dt_sigma_t))
  pit_dt_t[res$testdata[,"robs"]>0, 1] <- pit_dt_t[res$testdata[,"robs"]>0, 2]

  ## distforest
  if(FALSE){
    # in sample
    pdf <- predict(res$df, type = "parameter")
    df_mu_l <- pdf$mu
    df_sigma_l <- pdf$sigma
    pit_df_l <- cbind(0, pnorm(res$learndata[,"robs"], mean = df_mu_l, sd = df_sigma_l))
    pit_df_l[res$learndata[,"robs"]>0, 1] <- pit_df_l[res$learndata[,"robs"]>0, 2]
  }
  # out of sample
  pdf <- predict(res$df, type = "parameter", newdata = res$testdata)
  df_mu_t <- pdf$mu
  df_sigma_t <- pdf$sigma
  pit_df_t <- cbind(0, pnorm(res$testdata[,"robs"], mean = df_mu_t, sd = df_sigma_t))
  pit_df_t[res$testdata[,"robs"]>0, 1] <- pit_df_t[res$testdata[,"robs"]>0, 2]

  ## gamlss
  g_learndata <- res$learndata
  g_learndata$robs <- Surv(g_learndata$robs, g_learndata$robs>0, type="left")
  if(FALSE){
    # in sample
    g_mu_l <- predict(res$g, what = "mu", type = "response")
    g_sigma_l <- predict(res$g, what = "sigma", type = "response")
    pit_g_l <- cbind(0, pnorm(res$learndata[,"robs"], mean = g_mu_l, sd = g_sigma_l))
    pit_g_l[res$learndata[,"robs"]>0, 1] <- pit_g_l[res$learndata[,"robs"]>0, 2]
  }  
  # out of sample
  g_mu_t <- predict(res$g, what = "mu", type = "response", newdata = res$testdata)
  g_sigma_t <- predict(res$g, what = "sigma", type = "response", newdata = res$testdata)
  pit_g_t <- cbind(0, pnorm(res$testdata[,"robs"], mean = g_mu_t, sd = g_sigma_t))
  pit_g_t[res$testdata[,"robs"]>0, 1] <- pit_g_t[res$testdata[,"robs"]>0, 2]

  #gamboostLSS
  if(FALSE){
    #in sample
    pgb <- predict(res$gb, parameter = c("mu","sigma"), type = "response")
    gb_mu_l <- pgb$mu
    gb_sigma_l <- pgb$sigma
    pit_gb_l <- cbind(0, pnorm(res$learndata[,"robs"], mean = gb_mu_l, sd = gb_sigma_l))
    pit_gb_l[res$learndata[,"robs"]>0, 1] <- pit_gb_l[res$learndata[,"robs"]>0, 2]
  }  
  # out of sample
  pgb <- predict(res$gb, parameter = c("mu","sigma"), type = "response", newdata = res$testdata)
  gb_mu_t <- pgb$mu
  gb_sigma_t <- pgb$sigma
  pit_gb_t <- cbind(0, pnorm(res$testdata[,"robs"], mean = gb_mu_t, sd = gb_sigma_t))
  pit_gb_t[res$testdata[,"robs"]>0, 1] <- pit_gb_t[res$testdata[,"robs"]>0, 2]

  #EMOS
  if(FALSE){
    # in sample
    ml_mu_l <- predict(res$ml, type = "location")     # returns parameter on response scale
    ml_sigma_l <- predict(res$ml, type = "scale")
    pit_ml_l <- cbind(0, pnorm(res$learndata[,"robs"], mean = ml_mu_l, sd = ml_sigma_l))
    pit_ml_l[res$learndata[,"robs"]>0, 1] <- pit_ml_l[res$learndata[,"robs"]>0, 2]
  }  
  # out of sample
  ml_mu_t <- predict(res$ml, type = "location", newdata = res$testdata)     # returns parameter on response scale
  ml_sigma_t <- predict(res$ml, type = "scale", newdata = res$testdata)
  pit_ml_t <- cbind(0, pnorm(res$testdata[,"robs"], mean = ml_mu_t, sd = ml_sigma_t))
  pit_ml_t[res$testdata[,"robs"]>0, 1] <- pit_ml_t[res$testdata[,"robs"]>0, 2]

@@

## Variable importance
if(file.exists("vimp_crps.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/vimp_crps.rda")
  load("vimp_crps.rda")
} else {

  # locally redefine logLik.distforest to use crps in varimp()
  logLik.distforest <- function(object, newdata = NULL, ...) {
    if(is.null(newdata)) {
      newdata <- object$data
    }

@@

  vimp_crps <- varimp(res$df, nperm = 1L)
  #vimp10_crps <- varimp(res$df, nperm = 10L)
  #vimp100_crps <- varimp(res$df, nperm = 100L)
  #vimp1000_crps <- varimp(res$df, nperm = 1000L)

  save(vimp_crps, file = "~/svn/partykit/pkg/disttree/inst/draft/vimp_crps.rda")
  #save(vimp10_crps, file = "~/svn/partykit/pkg/disttree/inst/draft/vimp10_crps.rda")
  #save(vimp100_crps, file = "~/svn/partykit/pkg/disttree/inst/draft/vimp100_crps.rda")
  #save(vimp1000_crps, file = "~/svn/partykit/pkg/disttree/inst/draft/vimp1000_crps.rda")

@@

\begin{figure}[t!]
\centering
\setkeys{Gin}{width=0.99\textwidth}
<<rain_axams_qqr_oosample, fig=TRUE, echo=FALSE, width = 6.5, pdf = FALSE, png = TRUE, resolution = 150>>=
# out of sample
set.seed(7)
par(mfrow = c(2, 2))
qqrplot(pit_df_t, nsim = 100, main = "Distributional forest", ylim = c(-5, 5), col = gray(0.04, alpha = 0.01), pch = 19)
qqrplot(pit_ml_t, nsim = 100, main = "EMOS",                  ylim = c(-5, 5), col = gray(0.04, alpha = 0.01), pch = 19)
qqrplot(pit_g_t,  nsim = 100, main = "Prespecified GAMLSS",   ylim = c(-5, 5), col = gray(0.04, alpha = 0.01), pch = 19)
qqrplot(pit_gb_t, nsim = 100, main = "Boosted GAMLSS",        ylim = c(-5, 5), col = gray(0.04, alpha = 0.01), pch = 19)
@
\caption{\label{fig:axams-qq} Out-of-sample residual QQ~plots (2009--2012) 
for station Axams based on models learned on data from 1985--2008.}
\end{figure}

@@

<<echo=FALSE, results=hide>>=
#### cross validation rain
if(file.exists("rain_Axams_7x10.rda")){
  load("rain_Axams_7x10.rda")
} else {
  source("rain_cross_7x10.R")
  library("gamlss.cens")
  gen.cens("NO", type = "left")
  rainres <- rain_cross(stationname = "Axams", 
                        seedconst = 7, ntree = 100,
                        forest_mtry = 27,
                        gamboost_cvr = TRUE,
                        frac = FALSE)

  save(rainres, file = "rain_Axams_7x10.rda")
}

rain_rmse <- matrix(0, ncol = ncol(rainres[[1]]$rmse), length(rainres)-1)
rain_ll <- matrix(0, ncol = ncol(rainres[[1]]$ll), length(rainres)-1)
rain_crps <- matrix(0, ncol = ncol(rainres[[1]]$crps), length(rainres)-1)

for(i in 1:(length(rainres)-1)){
  rain_rmse[i,] <- colMeans(rainres[[i]]$rmse, na.rm = TRUE)
  rain_ll[i,] <- colMeans(rainres[[i]]$ll, na.rm = TRUE)
  rain_crps[i,] <- colMeans(rainres[[i]]$crps, na.rm = TRUE)
}

colnames(rain_rmse) <- colnames(rain_ll) <- colnames(rain_crps) <- colnames(rainres[[1]]$rmse)
@

\begin{figure}[t!]
\centering
<<rain_cross_axams_crps_skill_score, fig=TRUE, echo=FALSE, height=4.5, width = 7>>=
boxplot(1 - rain_crps[,c(2,3,4)] / rain_crps[,6], ylim = c(-0.005, 0.065),
        names = c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS"),
        ylab = "CRPS skill score", col = "lightgray") 
abline(h = 0, col = pal["EMOS"], lwd = 2)
@
\caption{\label{fig:axams-crps}CRPS skill score from the 10 times 
7-fold cross validation at station Axams (1985--2012). The horizontal orange 
line pertains to the reference model EMOS.}
\end{figure}

@@

<<echo = FALSE>>=
#### prediction over all stations 24 - 4
if(file.exists("rain_pred_24to4.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/rain_pred_24to4.rda")
  load("rain_pred_24to4.rda")
} else {

  source("rain_pred_24to4.R")
  library("gamlss.cens")
  gen.cens("NO", type = "left")

  res <- rain_pred(seedconst = 7,
                   gamboost_cvr = TRUE,
                   frac = FALSE)

  save(res, file = "rain_pred_24to4.rda")
}

#ll_all <- res[[1]]$results["ll",]
#for(i in 2:(length(res)-1)) ll_all <- rbind(ll_all, res[[i]]$results["ll",])

#rmse_all <- res[[1]]$results["rmse",]
#for(i in 2:(length(res)-1)) rmse_all <- rbind(rmse_all, res[[i]]$results["rmse",])

crps_all <- res[[1]]$results["crps",]
for(i in 2:(length(res)-1))  crps_all <- rbind(crps_all, res[[i]]$results["crps",])

#colnames(ll_all) <- colnames(rmse_all) <- 
  colnames(crps_all) <- colnames(res[[1]]$results)

# skill score
s <- 1 - crps_all[, 2:4]/crps_all[,6]
colnames(s) <- c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS")

@@

load("plot_map_rain/demo/tirol.gadm.rda")
load("plot_map_rain/demo/tirol.dem.rda")
load("plot_map_rain/demo/ehyd.statlist.rda")

# Create SpatialPointsDataFrame from station list
sp <- SpatialPointsDataFrame(subset(ehyd.statlist[res$complete_stations,],
                                    select=c(lon,lat)),
                             data = subset(ehyd.statlist[res$complete_stations,],
                                           select = -c(lon,lat)),
                             proj4string = crs(tirol.dem))
@

@@

  layout(cbind(1, 2), width = c(9, 1))
  par(mar = c(5,4,4,0.1))
  raster::image(tirol.dem, col = rev(gray.colors(100)),
                main="Stations in Tyrol", ylab = "Latitude", xlab = "Longitude", 
                xlim = c(9.8,13.2), 
                ylim = c(46.6, 47.87))
  plot(tirol.gadm, add = TRUE)
  points(sp, pch = c(21, 24, 25, 22)[bst], bg = clr[grp], col = "black", las = 1, cex = 1.5)
  legend(x = 9.8, y = 47.815, pch = c(21, 24, 25, 22), legend = c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS", "EMOS"), cex = 1, bty = "n")
  text(x = 10.3, y = 47.82, labels = "Models with lowest CRPS")
  mtext("CRPS\ndifference", side=4, las = TRUE, at = c(x = 13.5, y = 47.76), line = 0.3)
  par(mar = c(0.5,0.2,0.5,2.3))
  ## legend
  plot(0, 0, type = "n", axes = FALSE, xlab = "", ylab = "",
       xlim = c(0, 1), ylim = c(-0.2, 0.2), xaxs = "i", yaxs = "i")
  rect(0, brk[-6], 0.5, brk[-1], col = rev(clr))
  axis(4, at = brk, las = 1, mgp=c(0,-0.5,-1))

@
\caption{\label{fig:all-map}Map of Tyrol coding the best-performing model
for each station (type of symbol). The color codes whether the distributional 
forest had higher (green) or lower (red) CRPS compared to the best of the 
other three models. The background shows the local topography, the darker the
higher \citep{srtm}.}
\end{figure}

@@

par(mfrow = c(2,2))
pithist(pit_df_t, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "Distributional forest", ylim = c(0,1.5))
pithist(pit_ml_t, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "EMOS", ylim = c(0,1.5))
pithist(pit_g_t, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "Prespecified GAMLSS", ylim = c(0,1.5))
pithist(pit_gb_t, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "Boosted GAMLSS", ylim = c(0,1.5))

>>>>

<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)
library("disttree")
library("crch")
library("countreg")
library("gamlss")
library("gamlss.cens")
gen.cens(NO, type = "left")
library("RainTyrol")

@@

<<echo=FALSE, results=hide>>=
#### Axams prediction 24 - 4
if(file.exists("Axams_24to4.rda")){
  load("Axams_24to4.rda")
} else {
  source("Axams_24to4.R")
  Axams_24to4 <- Axams_24to4()
  save(Axams_24to4, file = "Axams_24to4.rda")
}

@@

pdays <- if(pday<19) c(pday, pday + 31, pday + 62, pday + 92) else c(pday, pday + 31, pday + 61, pday + 92)
pdf <- predict(Axams_24to4$df, newdata = Axams_24to4$testdata[pdays,], type = "parameter")
df_mu <- pdf$mu
df_sigma <- pdf$sigma
df_exp <- pnorm(df_mu/df_sigma) * (df_mu + df_sigma * (dnorm(df_mu/df_sigma) / pnorm(df_mu/df_sigma)))
cbind(df_exp, Axams_24to4$testdata[pdays,"robs"])

# plot predicted distributions together with observations
set.seed(7)
x <- c(0.01, sort(runif(500,0.01,8)))
y1 <- crch::dcnorm(x, mean = df_mu[1], sd = df_sigma[1], left = 0)
y2 <- crch::dcnorm(x, mean = df_mu[2], sd = df_sigma[2], left = 0)
y3 <- crch::dcnorm(x, mean = df_mu[3], sd = df_sigma[3], left = 0)
y4 <- crch::dcnorm(x, mean = df_mu[4], sd = df_sigma[4], left = 0)
dayending <- if(pday > 3) "th" else switch(pday, "1" = {dayending <- "st"}, "2" = {dayending <- "nd"}, "3" = {dayending <- "rd"})

@@

# predictions
pred1 <- c(Axams_24to4$testdata[pdays,"robs"][1], crch::dcnorm(Axams_24to4$testdata[pdays,"robs"][1], mean = df_mu[1], sd = df_sigma[1], left = 0))
pred2 <- c(Axams_24to4$testdata[pdays,"robs"][2], crch::dcnorm(Axams_24to4$testdata[pdays,"robs"][2], mean = df_mu[2], sd = df_sigma[2], left = 0))
pred3 <- c(Axams_24to4$testdata[pdays,"robs"][3], crch::dcnorm(Axams_24to4$testdata[pdays,"robs"][3], mean = df_mu[3], sd = df_sigma[3], left = 0))
pred4 <- c(Axams_24to4$testdata[pdays,"robs"][4], crch::dcnorm(Axams_24to4$testdata[pdays,"robs"][4], mean = df_mu[4], sd = df_sigma[4], left = 0))

@@

############
# get predicted parameter of all models for testdata

# distributional tree
pdt <- predict(Axams_24to4$dt, newdata = Axams_24to4$testdata, type = "parameter")
dt_mu <- pdt$mu
dt_sigma <- pdt$sigma
dt_exp <- pnorm(dt_mu/dt_sigma) * (dt_mu + dt_sigma * (dnorm(dt_mu/dt_sigma) / pnorm(dt_mu/dt_sigma)))

# distributional forest
pdf <- predict(Axams_24to4$df, newdata = Axams_24to4$testdata, type = "parameter")
df_mu <- pdf$mu
df_sigma <- pdf$sigma
df_exp <- pnorm(df_mu/df_sigma) * (df_mu + df_sigma * (dnorm(df_mu/df_sigma) / pnorm(df_mu/df_sigma)))

# prespecified GAM
g_mu <- predict(Axams_24to4$g, newdata = Axams_24to4$testdata, what = "mu", type = "response", data = Axams_24to4$g_learndata)
g_sigma <- predict(Axams_24to4$g, newdata = Axams_24to4$testdata, what = "sigma", type = "response", data = Axams_24to4$g_learndata)
g_exp <- pnorm(g_mu/g_sigma) * (g_mu + g_sigma * (dnorm(g_mu/g_sigma) / pnorm(g_mu/g_sigma)))

# boosted GAM
pgb <- predict(Axams_24to4$gb, newdata = Axams_24to4$testdata, parameter = c("mu","sigma"), type = "response")
gb_mu <- pgb$mu
gb_sigma <- pgb$sigma
gb_exp <- pnorm(gb_mu/gb_sigma) * (gb_mu + gb_sigma * (dnorm(gb_mu/gb_sigma) / pnorm(gb_mu/gb_sigma)))

# EMOS
ml_mu <- predict(Axams_24to4$ml, type = "location", newdata = Axams_24to4$testdata)
ml_sigma <- predict(Axams_24to4$ml, type = "scale", newdata = Axams_24to4$testdata)
ml_exp <- pnorm(ml_mu/ml_sigma) * (ml_mu + ml_sigma * (dnorm(ml_mu/ml_sigma) / pnorm(ml_mu/ml_sigma)))

############
## PIT histograms and qqrplots

## get predicted parameter
# for learndata and testdata
{
  set.seed(7)

  # distributional tree
  pit_dt <- cbind(0, pnorm(Axams_24to4$testdata[,"robs"], mean = dt_mu, sd = dt_sigma))
  pit_dt[Axams_24to4$testdata[,"robs"]>0, 1] <- pit_dt[Axams_24to4$testdata[,"robs"]>0, 2]

  # distributional forest
  pit_df <- cbind(0, pnorm(Axams_24to4$testdata[,"robs"], mean = df_mu, sd = df_sigma))
  pit_df[Axams_24to4$testdata[,"robs"]>0, 1] <- pit_df[Axams_24to4$testdata[,"robs"]>0, 2]

  # prespecified GAM
  pit_g <- cbind(0, pnorm(Axams_24to4$testdata[,"robs"], mean = g_mu, sd = g_sigma))
  pit_g[Axams_24to4$testdata[,"robs"]>0, 1] <- pit_g[Axams_24to4$testdata[,"robs"]>0, 2]

  # boosted GAM
  pit_gb <- cbind(0, pnorm(Axams_24to4$testdata[,"robs"], mean = gb_mu, sd = gb_sigma))
  pit_gb[Axams_24to4$testdata[,"robs"]>0, 1] <- pit_gb[Axams_24to4$testdata[,"robs"]>0, 2]

  # EMOS
  pit_ml <- cbind(0, pnorm(Axams_24to4$testdata[,"robs"], mean = ml_mu, sd = ml_sigma))
  pit_ml[Axams_24to4$testdata[,"robs"]>0, 1] <- pit_ml[Axams_24to4$testdata[,"robs"]>0, 2]

@@

## Variable importance
if(file.exists("vimp_crps.rda")){
  load("vimp_crps.rda")
} else {

  set.seed(7)
  # locally redefine logLik.distforest to use crps in varimp()
  logLik.distforest <- function(object, newdata = NULL, ...) {
    if(is.null(newdata)) {
      newdata <- object$data
    }

@@

  vimp_crps <- varimp(Axams_24to4$df, nperm = 1L)

  save(vimp_crps, file = "vimp_crps.rda")

@@

\begin{figure}[t!]
\centering
\setkeys{Gin}{width=0.99\textwidth}
<<rain_axams_qqr_oosample, fig=TRUE, echo=FALSE, width = 6.5, pdf = FALSE, png = TRUE, resolution = 150>>=
# out of sample
set.seed(7)
par(mfrow = c(2, 2))
qqrplot(pit_df, nsim = 100, main = "Distributional forest", ylim = c(-5, 5), col = gray(0.04, alpha = 0.01), pch = 19)
qqrplot(pit_ml, nsim = 100, main = "EMOS",                  ylim = c(-5, 5), col = gray(0.04, alpha = 0.01), pch = 19)
qqrplot(pit_g,  nsim = 100, main = "Prespecified GAMLSS",   ylim = c(-5, 5), col = gray(0.04, alpha = 0.01), pch = 19)
qqrplot(pit_gb, nsim = 100, main = "Boosted GAMLSS",        ylim = c(-5, 5), col = gray(0.04, alpha = 0.01), pch = 19)
@
\caption{\label{fig:axams-qq} Out-of-sample residual QQ~plots (2009--2012) 
for station Axams based on models learned on data from 1985--2008.}
\end{figure}

@@

<<echo=FALSE, results=hide>>=
#### cross validation rain
if(file.exists("crps_cross.rda")){
  load("crps_cross.rda")
} else {

  nrep_cross <- 10
  seed <- 7

  res_cross <- mclapply(1:nrep_cross,
                        function(i){

                          set.seed(seed*i)

                          # randomly split data in 7 parts each including 4 years
                          years <- 1985:2012
                          testyears <- list()
                          for(j in 1:7){
                            testyears[[j]] <- sample(years, 4, replace = FALSE)
                            years <- years[!(years %in% testyears[[j]])]
                          }

                          #crps <- matrix(nrow = 7, ncol = 7)
                          reslist <- list()
                          for(k in 1:7){
                            test <- testyears[[k]]
                            train <- c(1985:2012)[!c(1985:2012) %in% test]

                            res <- eval(station = "Axams",
                                        train = train,
                                        test = test,
                                        gamboost_cvr = TRUE)

                            #crps[k,] <- res$crps
                            reslist[[k]] <- res
                          }

                          #colnames(crps) <- names(res$crps)
                          return(reslist)
                        },
                        mc.cores = detectCores() - 1
  )

  # extract CRPS
  crps_cross <- matrix(nrow = nrep_cross, ncol = 7)
  # loop over all repetitions
  for(i in 1:length(res_cross)){
    #loop over all 7 folds (for 7 methods)
    crps_cross_int <- matrix(nrow = length(res_cross[[1]]), ncol = 7)
    for(j in 1:length(res_cross[[1]])){
      crps_cross_int[j,] <- res_cross[[i]][[j]]$crps
    }
    crps_cross[i,] <- colMeans(crps_cross_int, na.rm = TRUE)
  }
  colnames(crps_cross) <- names(res_cross[[1]][[1]]$crps)

  save(crps_cross, file = "crps_cross.rda")
  save(res_cross, file = "res_cross.rda")

}
@

\begin{figure}[t!]
\centering
<<rain_cross_axams_crps_skill_score, fig=TRUE, echo=FALSE, height=4.5, width = 7>>=
boxplot(1 - crps_cross[,c(2,3,4)] / crps_cross[,6], ylim = c(-0.005, 0.065),
        names = c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS"),
        ylab = "CRPS skill score", col = "lightgray") 
abline(h = 0, col = pal["EMOS"], lwd = 2)
@
\caption{\label{fig:axams-crps}CRPS skill score from the 10 times 
7-fold cross validation at station Axams (1985--2012). The horizontal orange 
line pertains to the reference model EMOS.}
\end{figure}

@@

<<echo = FALSE>>=
#### prediction over all stations 24 - 4
if(file.exists("crps_24to4_all.rda")){
  load("crps_24to4_all.rda")
} else {

  data("StationsTyrol")
  stations <- StationsTyrol$name
  test <- 2009:2012
  train <- 1985:2008

  res_24to4_all <- mclapply(1:length(stations),
                            function(i){

                              set.seed(7)

                              res <- eval(station = stations[i],
                                          train = train,
                                          test = test,
                                          gamboost_cvr = TRUE)

                              return(res)
                            },
                            mc.cores = detectCores() - 1
  )

  # extract crps
  crps_24to4_all <- matrix(nrow = length(stations), ncol = 7)
  # loop over all stations
  for(i in 1:length(stations)){
    crps_24to4_all[i,] <- res_24to4_all[[i]]$crps
  }

  colnames(crps_24to4_all) <- names(res_24to4_all[[1]]$crps)
  rownames(crps_24to4_all) <- stations

  save(crps_24to4_all, file = "crps_24to4_all.rda")
  save(res_24to4_all, file = "res_24to4_all.rda")

# skill score
s <- 1 - crps_24to4_all[, 2:4]/crps_24to4_all[,6] 
colnames(s) <- c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS")

@@

load("plot_map_rain/data/tirol.rda")
load("plot_map_rain/data/dem.rda")

data(StationsTyrol)
# Create SpatialPointsDataFrame from station list
sp <- SpatialPointsDataFrame(subset(StationsTyrol,
                                    select=c(lon,lat)),
                             data = subset(StationsTyrol,
                                           select = -c(lon,lat)),
                             proj4string = crs(dem))
@

@@

  ## plot map of Tyrol with all 95 observations
  layout(cbind(1, 2), width = c(9, 1))
  par(mar = c(5,4,4,0.1))
  raster::image(dem, col = rev(gray.colors(100)),
                main="Stations in Tyrol", ylab = "Latitude", xlab = "Longitude", 
                xlim = c(9.8,13.2), 
                ylim = c(46.6, 47.87))
  plot(tirol, add = TRUE)
  points(sp, pch = c(21, 24, 25, 22)[bst], bg = clr[grp], col = "black", las = 1, cex = 1.5)
  legend(x = 9.8, y = 47.815, pch = c(21, 24, 25, 22), legend = c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS", "EMOS"), cex = 1, bty = "n")
  text(x = 10.3, y = 47.82, labels = "Models with lowest CRPS")
  mtext("CRPS\ndifference", side=4, las = TRUE, at = c(x = 13.5, y = 47.76), line = 0.3)
  par(mar = c(0.5,0.2,0.5,2.3))
  ## legend
  plot(0, 0, type = "n", axes = FALSE, xlab = "", ylab = "",
       xlim = c(0, 1), ylim = c(-0.2, 0.2), xaxs = "i", yaxs = "i")
  rect(0, brk[-6], 0.5, brk[-1], col = rev(clr))
  axis(4, at = brk, las = 1, mgp=c(0,-0.5,-1))

@
\caption{\label{fig:all-map}Map of Tyrol coding the best-performing model
for each station (type of symbol). The color codes whether the distributional 
forest had higher (green) or lower (red) CRPS compared to the best of the 
other three models. The background shows the local topography \citep{Robinson:2014}.}
\end{figure}

@@

par(mfrow = c(2,2))
pithist(pit_df, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "Distributional forest", ylim = c(0,1.5))
pithist(pit_ml, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "EMOS", ylim = c(0,1.5))
pithist(pit_g, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "Prespecified GAMLSS", ylim = c(0,1.5))
pithist(pit_gb, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "Boosted GAMLSS", ylim = c(0,1.5))
